---
title: "Climate Hazards for Hapori Māori"
output:
  html_document:
    theme: flatly
    includes:
      after_body:
        - script/index_script.js
        - script/script.js
---
<script>
  var leafletMap;
</script>
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
paper_link <-
  paste0(
    "https://www.tpk.govt.nz",
    "/en/o-matou-mohiotanga",
    "/climate/understanding-climate-hazards-for-hapori-maori-ins"
  )
guide_link <-
  paste0(
    "https://camo.githubusercontent.com",
    "/b26592afe00593d868948da3311b553d3c40fd6922b3875a51b31e7123baacce",
    "/68747470733a2f2f7069766f747461626c652e6a732e6f72672f696d616765732f616e696d6174696f6e2e676966"
  )
```

<div id="headline" style="color:#007C9D;font-size:medium;margin-bottom: 21px;"> <em>
    Climate change presents significant challenges that necessitate robust
    policies and implementable strategies to support resilience and avoid
    maladaptation. The first step towards effective policy interventions is
    gaining a comprehensive understanding of the risks faced by our communities.
    By tailoring adaptation actions to address these risks, we can better serve
    our communities.
    </em>
</div>

<p id="introbody">
We have embarked on an initial exploratory analysis to examine what risks exist
specifically for hapori Māori. By assessing the socio-economic
circumstances and resilience of hapori Māori exposed to various climate hazards,
the findings published in <a href=`r paper_link` target="_blank">this</a>
report offer valuable insights to inform the immediate shaping of future climate
change policy design.

This page includes key visualisations from this study. We have included a
comprehensive map of the climate hazards projected to affect the greatest
number of Māori households in the future, namely heatwaves, droughts, extreme
hot days, wet spells, extreme rainfall and sea level rise. It also covers the
hazards already significantly impacting Māori households, such as flooding.
Furthermore, the page delves into the socio-economic circumstances and
resilience of Māori communities exposed to these climate hazards, shedding
light on disparities and areas that may require targeted policies. For instance,
it emphasises the vulnerability of hapori Māori in coastal areas due to sea 
level rise, as well as the challenges faced by older Māori households with lower
adaptability due to factors like social isolation and lack of support systems.

<br> Before using this application, please take a moment to read the Methodology
and Data section. Click on the following tabs to see the results: <br>
</p>

### <span></span> {.tabset}

#### Maps
<div class="row">
<div class="col-md-3">
This map provides both overall climate and socio-economic information at a
regional level. You can use the layers control icon to switch between these two
layers on the map. Additionally, we have included more detailed maps for each
regional council at the Statistical Area 2 (SA2) level. These maps can be
accessed via popup windows by clicking on a region<span class="mapDesc"> or
using the navbar at the top of this page</span>. The climate-based maps for SA2s
include various socio-economic filter layers that can be overlaid on climate
information. Similarly, the socio-economic-based maps can be masked by climate
filter layers. Note that a higher decile number corresponds to a greater risk
level in these maps.

<h5 style="color:#007C9D">How to use:</h5>
* Zoom in/out using the buttons or the mouse scroll wheel, move by dragging
* Fullscreen and reset view buttons
* Download button to export the map with current view to a PNG file
* Layer control for base (and overlay) maps, and toggling marker locations
(marae)
* Search button/toolbar based on SA2 names
* Hover and popup information for shapes and markers
</div>
<div class="col-md-9" id="the_map">
```{r out.width='calc(100% + 15px)', out.height='750px', echo=FALSE}
library(leaflet)
readRDS("../../processed_data/national_maps/overall_base_widget.rds")
```
</div>
</div>

#### Tables
<div class="row">
<div class="col-md-3" style="padding-top:12px;">
This visualisation tool facilitates data exploration and analysis by
transforming the dataset into a concise summary table. Additionally, it offers
an optional 2-dimensional drag-and-drop user interface that allows users to
manipulate this summary table, effectively converting it into a pivot table.
Furthermore, we have included several plugins that allow the summary table to
be rendered as various types of charts, effectively turning the pivot table
into a dynamic pivot chart. The information provided in this table excludes
SA2s with missing climate or socio-economic information. Note that the table
displays output for attribute values that exist in the data. If there is no
data related to a specific category, the corresponding cells are left blank.

<h5 style="color:#007C9D">How to use:</h5>
* From the top-left dropdown menu, select a plugin such as `Table` or
`Bar Chart`
* Next, choose an aggregator method from the dropdown menu. Options include
`Count`, `Sum`, and more
* Drag and drop the desired attributes into either the row or column container
* Filter the attributes by categories
* Arrange the results either by count or alphabetically
* Click <a href=`r guide_link` target="_blank">here</a> for a visualised guide
</div>
<div class="col-md-9" id="the_table"
     style="height:762px;overflow:auto;padding-top:8px;padding-left:11px;">
```{r out.width='100%', out.height='762px', echo=FALSE}
htmltools::includeHTML("../../output_data/pivot_table/pivot_table.html")
```
</div>
</div>

#### Location Insights
<div class="row">
<div class="col-md-3">
This tool is designed to help users search for specific locations and gain
comprehensive insights about them. By simply entering the name of a city, town,
or suburb into the search bar, users can access detailed information including
geographic data, demographic statistics, climate variables, and socio-economic
indicators of the overlapping SA2 areas. The findings provide
a comprehensive view of the area, making it easier for users to understand the
characteristics and conditions of different locations. The location data is
based on New Zealand Suburbs and Localities, a dataset administered by Toitū Te
Whenua Land Information New Zealand (last updated on August 2024).

<h5 style="color:#007C9D">How to use:</h5>
* Enter the name of a place into the search bar and select the most relevant
suggestion
* The search process includes both main and alternative names, but the
alternative names are not displayed in the suggestions
* Select `All Results` to view all overlapping SA2 areas, and select an SA2 to
expand its information, the search results are ordered by the extent to which a
place's shape is covered by an SA2
* Select either the `Climate` or `SES` button to view the corresponding
visualisation 
</div>
<div class="col-md-9" id="the_finder"
     style="height:750px;padding-top:0px;padding-left:14px;padding-right:0px;">
```{r out.width='100%', out.height='762px', echo=FALSE}
htmltools::includeHTML("../../output_data/finder/finder.html")
```
</div>
</div>
