---
title: "Climate Hazards for Hapori Māori"
---

```{js, echo=FALSE}

$('.title').css('color', '#007C9D');
$('.tabset').css('color', '#007C9D');
$(document).ready(function() {
    $head = $('#navbar');
    $head.prepend('<img src=\"tpk_logo.svg\" style=\"height:40px;top:5px;right:15px;position:absolute;\"/>');
    setTimeout(function(){
      var localLinks = document.querySelectorAll('a[href^="maps');
      for(var i = 0; i < localLinks.length; i++) {
        localLinks[i].setAttribute('target', '_blank');
      }
    }, 20);
    document.querySelector(".rpivotTable").setAttribute('style', 'position:relative;');
});
```

<br>
<span style="color:#007C9D;font-size:medium"> <em>
Climate change presents significant challenges that necessitate robust policies
and implementable strategies to support resilience and avoid maladaptation.
The first step towards effective policy interventions is gaining a
comprehensive understanding of the risks faced by our communities. By tailoring
adaptation actions to address these risks, we can better serve our communities.
</em></span>
<br><br>

We have embarked on an initial exploratory analysis to examine what risks exist
specifically for hapori Māori. By assessing the socio-economic
circumstances and resilience of hapori Māori exposed to various climate hazards,
the findings presented in this report offer valuable insights to inform the
immediate shaping of future climate change policy design.
This page includes key visualisations from this study. We have included a
comprehensive map of the climate hazards projected to affect the greatest
number of Māori households in the future, namely heatwaves, droughts, extreme
hot days, wet spells, extreme rainfall and sea level rise. It also covers the
hazards already significantly impacting Māori households, such as flooding.
Furthermore, the page delves into the socio-economic circumstances and
resilience of Māori communities exposed to these climate hazards, shedding
light on disparities and areas that may require targeted policies. For instance,
it emphasizes the vulnerability of hapori Māori in coastal areas due to sea 
level rise, as well as the challenges faced by older Māori households with lower
adaptability due to factors like social isolation and lack of support systems.

<br>
Click on the following tabs to see the results:
<br>

### <span></span> {.tabset}

#### Maps

:::::: {.columns}
::: {.column width="25%"}
These maps are based on the climate datasets provided by the National Institute
of Water and Atmospheric Research <a href="https://niwa.co.nz/" target="_blank">(NIWA)</a>.
We obtained climate maps for a
baseline period from 1986 to 2005 and projected changes averaged over time.
The projections for this analysis were based on a low-moderate global emission
trajectory called Representative Concentration Pathways (RCP), developed by
IPCC and used in global climate models.

For flooding, we obtained the 2021 composite flood hazard area maps from NIWA,
with consent from district and regional councils. These were created from
modelled and historic flood hazard maps and flood prone soil maps, publicly
available in August 2018.

These maps also include various socio-economic filter layers that can be
overlaid on climate information. To construct household-level measures of
socio-economic circumstances, we utilized the Integrated Data Infrastructure
<a href="https://www.stats.govt.nz/integrated-data/integrated-data-infrastructure/" target="_blank">(IDI)</a>.
The social indicators were selected based on their ability to capture
the resilience and adaptive capacity of hapori Māori. These indicators included
variables such as income, employment, education, health, and housing.
Statistical techniques and feedback from subject matter experts were used to
consolidate the numerous social indicators (approximately 100) into composite
indices representing different aspects of socio-economic risk.

<h5 style="color:#007C9D">How to use:</h5>
* Zoom in/out using the buttons or the mouse scroll wheel, move by dragging
* Fullscreen and reset view buttons
* Download button to export the map with current view to a PNG file
* Layer control for base and overlay maps, and toggling marker locations (marae)
* Search button/toolbar based on SA2 names
* Detect location button to highlight user’s location (permission required)
* Hover and popup information for shapes and markers

:::
::: {.column width="2%"}

:::
::: {.column width="73%"}
```{r out.width='100%', out.height='750px', echo=FALSE}
library(leaflet)
readRDS("../../processed_data/national_maps/overall_base_widget.rds")
```
:::
::::::

#### Tables
:::::: {.columns}
::: {.column width="25%"}

This visualisation enables data exploration and analysis by turning the project's dataset
into a summary table and then optionally adding a true 2-d drag'n'drop UI to
allow users to manipulate this summary table, turning it into a pivot table,
very similar to the one found in older versions of Microsoft Excel with a bunch
of extra developer-oriented features and some visualization effects.
We have also included several add-ons, with which the summary table can be
rendered as various kinds of charts, turning the pivot table into a pivot chart.

<h5 style="color:#007C9D">How to use:</h5>
* Select a plugin from top left dropdown menu, e.g., Table, Bar Chart, etc.
* Select an aggregator method from the next dropdown menu, e.g., Count, Integer Sum, etc.
* Drag'n'drop desired attributes in a row or column container
* Filter by categories for each of the attributes 
* Sort the results by count or alphabetically
* Click <a href="https://camo.githubusercontent.com/b26592afe00593d868948da3311b553d3c40fd6922b3875a51b31e7123baacce/68747470733a2f2f7069766f747461626c652e6a732e6f72672f696d616765732f616e696d6174696f6e2e676966" target="_blank">here</a> for a visualised guide 

:::
::: {.column width="2%"}
\
<!-- an empty Div (with a white space), serving as a column separator -->
:::
::: {.column style="width:73%;overflow:scroll;padding-top:7px"}
```{r, out.width='100%', out.height='750px', echo=FALSE}
library(rpivotTable)
readRDS("../../processed_data/pivot_table/pivot_widget.rds")
```
:::
::::::
